Concurrent Programming with Harmony

### 2020-07-30 book/paper.tex
Latest

Either way, it vacates one of the semaphores next.
\texttt{V\_one()} is the baton passing function that selects which of the
binary semaphores to vacate.
Method \texttt{V\_one()} first checks to see if there are any readers or
writers waiting.  If there are readers waiting and there are no writers
in the critical section, it vacates the semaphore associated with
readers waiting.  If there are writers waiing and there are no readers
nor writers in the critical section, it vacates the semaphore associated
with writers waiting.  If both conditions hold, it could vacate either
one---it would not matter.  However, it can only vacate one of the semaphores.
If none of the conditions hold, \texttt{V\_one()} vacates \textit{mutex}.

Let us return now to \texttt{rlock\_acquire}. In the case that the process,
say $p$,
found there is a writer in the critical section, it vacates
\textit{mutex} and procures \textit{r\_sema},
blocking on the semaphore associated
with waiting readers.  When later some other process passes the baton to
$p$ by incrementing \textit{r\_sema}, $p$ updates the state by decrementing
\textit{r\_waiting} and incrementing \textit{r\_entered}.  $p$
then invokes \texttt{V\_one()} one more time.

As an example, consider the case where there is a writer in the critical
section and there are two readers waiting.  Let us see what happens when
the writer calls \texttt{wlock\_release}:
\begin{enumerate}
\item After procuring \textit{mutex},
the sum of the three semaphores is 0.  The writer then decrements
\textit{w\_entered} and calls \texttt{V\_one()}.
\item \texttt{V\_one()} finds that there are no writers in the critical section
and there are two readers waiting.  It therefore vacates \textit{r\_sema}.
The sum of the semaphores is now 1, because \textit{r\_sema} is 1.  This
guarantees that only a waiting reader, procuring \textit{r\_sema}, can enter
the critical section.
\item When it does, the reader decrements \textit{r\_waiting}
from 2 to 1, and increments \textit{r\_entered} from 0 to 1.
The reader finally calls \texttt{V\_one()}.
\item Again, \texttt{V\_one()} finds that there are no writers and
that there are readers waiting, so again it vacates \textit{r\_sema},
passing the baton to the one remaining waiting reader.
\item The remaining reader decrements \textit{r\_waiting} from 1 to 0 and
increments \textit{r\_entered} from 1 to 2.
\item Finally, the remaining reader  calls \texttt{V\_one()}.
\texttt{V\_one()} does not find any processes waiting,
and so it vacates \textit{mutex}.
\end{enumerate}

>>>>

Either way, it vacates one of the semaphores next.
\texttt{V\_one()} is the baton passing function that selects which of the
binary semaphores to vacate.
Method \texttt{V\_one()} first checks to see if there are any readers or
writers waiting.  If there are readers waiting and there are no writers
in the critical section, it vacates the semaphore associated with
readers waiting, passing the baton to the next reader.
If there are writers waiting and there are no readers
nor writers in the critical section, it vacates the semaphore associated
with writers waiting, passing the baton to the next writer.
If both conditions hold, it could vacate either
one---it would not matter.
However, it can only vacate one of the semaphores.
In this implementation, it prefers to pass the baton to another reader.
If none of the conditions hold, \texttt{V\_one()} vacates \textit{mutex},
leaving the batom for a future process to pick up.

Let us return now to \texttt{rlock\_acquire}. In the case that the process,
say $p$,
found there is a writer in the critical section, it vacates
\textit{mutex} and procures \textit{r\_sema},
blocking on the semaphore associated
with waiting readers.  When later some other process passes the baton to
$p$ by incrementing \textit{r\_sema}, $p$ updates the state by decrementing
\textit{r\_waiting} and incrementing \textit{r\_entered}.  $p$
then invokes \texttt{V\_one()} to pass the baton to another waiting
reader, if any, or to leave it for another process to pick up.

As an example, consider the case where there is a writer in the critical
section and there are two readers waiting.  Let us see what happens when
the writer calls \texttt{wlock\_release}:
\begin{enumerate}
\item After procuring \textit{mutex},
the sum of the three semaphores is 0.  The writer then decrements
\textit{w\_entered} and calls \texttt{V\_one()}.
\item \texttt{V\_one()} finds that there are no writers in the critical section
and there are two readers waiting.  It therefore vacates \textit{r\_sema}.
The sum of the semaphores is now~1, because \textit{r\_sema} is~1.  This
guarantees that only a waiting reader, procuring \textit{r\_sema}, can enter
the critical section.
\item When it does, the reader decrements \textit{r\_waiting}
from 2 to 1, and increments \textit{r\_entered} from~0 to~1.
The reader finally calls \texttt{V\_one()}.
\item Again, \texttt{V\_one()} finds that there are no writers and
that there are readers waiting, so again it vacates \textit{r\_sema},
passing the baton to the one remaining waiting reader.
\item The remaining reader decrements \textit{r\_waiting} from 1 to 0 and
increments \textit{r\_entered} from~1 to~2.
\item Finally, the remaining reader  calls \texttt{V\_one()}.
\texttt{V\_one()} does not find any processes waiting,
and so it vacates \textit{mutex}.
\end{enumerate}
